Versatile Robust Clustering of Ad Hoc Cognitive Radio Network

### 2016-07-24 journal-draft/bare_jrnl_comsoc.tex
change one plot, and minor changes
the version before sending to James in July

\definecolor{slightgray}{gray}{.90} 
\usepackage{rotating}
\usepackage{hhline}

@@

\newcommand{\eg}{e.g., }
\newcommand{\ie}{i.e., }
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
%\usepackage{algorithmic}
\usepackage{times}
%\usepackage{mathptmx}
\usepackage{mathtools}
\usepackage[draft,nomargin,marginclue,footnote,silent]{fixme}
\setcounter{tocdepth}{2}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
Cluster structure in cognitive radio networks facilitates cooperative spectrum sensing, routing and other functionalities.
The availability of unlicensed channels which are available for every member in a cluster decides the robustness of that cluster against the licensed users' influence.
Thus there should be more unlicensed channels in the clusters.
In the process of forming clusters, every secondary user decides with whom to form a cluster, or which cluster to join.
Congestion game model is adopted to analyse this process, which not only contributes the algorithm design directly, but also provides guarantee of convergence into Nash Equilibrium and convergence speed.
%
Our proposed distributed clustering scheme outperforms the comparison scheme in terms of cluster robustness against the primary users, convergence speed and volume of control messages.
Furthermore, the proposed clustering solution is versatile to fulfil other requirements such like fast convergence and cluster size control.
Besides, we prove the clustering problem is NP-hard, and also propose the centralized solution.
The extensive simulation supports our claims.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Cognitive Radio, Cluster, Robust, game theory, congestion game, distributed, centralised, size control.
\end{IEEEkeywords}

@@

\section{Introduction}
\label{intro}
%CR concept
\IEEEPARstart{C}{ognitive} radio (CR) is a promising technology to solve the spectrum scarcity problem~\cite{Mitola}.
Unlicensed users access the spectrum allocated to them whenever there is information to be transmitted.
In contrast, unlicensed users can only access the licensed spectrum after validating the channel is unoccupied by licensed users.
This refers to the process of sensing a particular channel and verifying (with a previously specified probability of error) it is not used by a primary user currently.
In this hierarchical spectrum access model~\cite{zhao_survey_DSA_2007}, the licensed users are also called primary users (PU), and the CR users are known as secondary users and constitute the cognitive radio networks (CRN).

%cluster
As to the operation of CRN, efficient spectrum sensing is identified to be critical to the success of cognitive radio networks~\cite{Sahai_FundamentalDesignTradeoffs2006}.
Cooperative spectrum sensing is able to effectively cope with noise uncertainty and channel fading, thus remarkably improves the sensing accuracy~\cite{coorperativeSensing_Akyildiz11}.
Collaborative sensing relies on the consensus of CR users within certain area, and decreases considerably the false sensing reports caused by fading and shadowing of reporting channel.
In this regard, clustering is regarded as an effective method in cooperative spectrum sensing~\cite{Sun07_clustering_spectrum_secsing, Zhao07}, as a cluster forms adjacent secondary users as a collectivity to perform spectrum sensing together.
%As a result, more accurate sensing result can be obtained by collaborative sensing, and the improvement on spectrum sensing decreases the interference originating from CR users to primary users, which is highly desirable.
%Also,  prevents CR users from using channels that are occupied by primary users. 
Clustering is also efficient to enable all CR devices within the same cluster to stop payload transmission on the operating channel and initiate the sensing process, so that all the CR users\footnote{The term $\textit{user}$ and $\textit{node}$ are used interchangeably in this paper, in particular, user is used when its networking or cognitive ability are discussed or stressed, and node is used when the network topology is discussed.} within the one cluster are able to vacate the channel swiftly when primary users are detected by at least one CR node residing in the cluster~\cite{willkomm08}.
With cluster structure, as CR users can be notified by cluster head (CH) or other cluster members about the possible collision, the possibility for them to interfere neighbouring clusters is reduced~\cite{centralizedSharing80222}. 
Clustering algorithm has also proposed to support routing in cognitive ad-hoc networks~\cite{Abbasi_survey_07}.

@@

%[robustness issue for clustering]
The communication within a cluster is conducted in the spectrum which is available for every member in that cluster.
Usually there are multiple unlicensed channels available for all the members in a cluster, which are referred as \textit{common control channels} (CCC).
When one or several members can not use one certain CCC because primary users are detected to appear on that channel, this channel will be excluded from the set of CCCs, in particular, if this channel is the working channel, then all the cluster members switch to another channel in the set of CCCs.
In the context of CRN, as the activity of primary users is controlled by licensed operators which are generally not known to CR users, the availability of CCCs for the formed clusters  is totally decided by primary users' activity.
In other words, the availability of CCCs for clusters is passive and can not guaranteed.
%
In CRN, one cluster survives the influence of primary users when at least one CCC is available for that cluster.
As the channel occupation by primary users is assumed to be uncontrollable to the CR users, a cluster formed with more CCCs will survive with higher probability.
Thus the number of CCCs in one cluster indicates robustness of it when facing ungovernable influence from primary users.
%It is easy to see that forming clusters with different neighbours leads to different amount of CCCs in the clusters.
As a result, how to form the clusters plays an important role on the robustness of clusters in CRN.

@@

To solely pursue cluster robustness against the primary users' activity, \ie to achieve more common channels within clusters, the ultimately best clustering strategy is ironically that each node constitutes one single node clusters.
%In that case, the common channels within cluster are the available channels available at that node's place.
Apparently this contradicts our motivation of proposing cluster in cognitive radio network.
This contradiction indicates that, the robustness discussed in terms of number of common channels carries little meaning when the sizes of formed clusters are not given consideration.
Besides, cluster size plays import roles in certain aspects.
For instance, cluster size is one decisive factor in power preservation~\cite{clustering_globecom11, EnergyEfficientClusteringRouting_2015}, and it also influences the accuracy of cooperative spectrum sensing~\cite{Consensus_based_clustering12}.
Hence, cluster size should be given consideration when discussing cluster robustness against primary users.

In this paper, a decentralized clustering approach ROSS (RObust Spectrum Sharing) is proposed to cover the issues of robustness and size control of clusters in CRN.
ROSS is able to form clusters with desired sizes, and the generated clusters are more robust than other clustering scheme which has claims on cluster robustness, \ie more secondary users residing in clusters against increasing influence from primary users.
Compared with previous work, ROSS involves much less control messages, and the generated clusters are significantly more robust.
%ROSS selects cluster head through coordination within its neighborhood, and then cluster membership is decided locally and its fast convergence is proved under game theoretic framework. 
%For our scheme we can prove convergence in cluster formation phase and resolve ambiguities with respect to cluster membership in a game-theoretic setting. 
We also propose the light weighted versions of ROSS, which involve less overheads and thus are more suitable for mobile networks.
Throughout this paper, we refer the clustering schemes on the basis of ROSS as \textit{variants of ROSS}, \ie the fast versions, or that with size control feature.

The rest of paper is organized as follows. 
After reviewing related work in section~\ref{related_work}, we present our system model in Section~\ref{sec:model}. 
Then we introduce our clustering scheme ROSS and its variants in section~\ref{ross}.
The clustering problem is given through analysis and a centralized scheme is proposed in section~\ref{centralized_scheme}.
Extensive performance evaluation is in section~\ref{performance}.
Finally, we conclude our work and point out direction future research in section~\ref{conclusion}.

@@

Prior to the emergence of open spectrum access, as an important method to manage network, clustering has been proposed in for ad hoc networks~\cite{Kawadia03,Lin97adaptiveclustering,Basagni99}, wireless mesh networks and sensor networks~\cite{Abbasi_survey_07}. 
In ad hoc and mesh networks, the major focus of clustering is to preserve connectivity (under static channel conditions) or to improve routing.
In sensor networks, the emphasis of clustering has been on longevity and coverage.
Overhead generated by clustering in ad hoc network is analysed in~\cite{clusterRoutingOverhead02infocom, clusterRoutingOverhead_wcnc04}.

As to cognitive radio networks, clustering schemes are also proposed, which target different aspects.
Work~\cite{Consensus_based_clustering12} improves spectrum sensing ability by grouping the CR users
with potentially best detection performance into the same cluster.
Clustering scheme~\cite{clustering_globecom11} obtains the best cluster size which minimizes power consumption caused by communication within and among clusters.
\cite{clustering_globecom11} proposes clustering strategy in cognitive radio network, which looks into the relationship between cluster size and power consumption and accordingly controlling the cluster size to decrease power consumption.
%The works~\cite{} introduced in Introduction section don't provide solution that how are the clusters formed.
%There have been several clustering schemes tailored for CRNs. 
Cogmesh is proposed in~\cite{Chen07} to construct clusters by the neighbour nodes which share local common channels, and by interacting with neighbour clusters, a mesh network in the context of open spectrum sharing is formed.
Robustness issue is not considered by this clustering approach.
\cite{TWC2012_cooperative_communication} targets on the QoS poisoning and energy efficiency. 
This approach first decides on the relay nodes which minimize transmission power consumption, then the chosen nodes become cluster heads and clusters are formed in a dynamic coalition process.
This work emphasis on power efficiency and doesn't take into account the channel availability and the issue of robustness of the formed clusters.
In~\cite{Zhao07, Affinity_clustering_09icccn}, the channel available to the largest set of one-hop neighbours is selected as common channel which yields a partition of the CRN into clusters. 
This approach minimizes the set of distinct frequency bands (and hence, the set of clusters) used as common channels within the CRN.
However, bigger cluster sizes generally lead to less options within one cluster to switch to if the common channel is reclaimed by a primary node. 
Hence, this scheme does not provide robustness to formed clusters. 
\cite{cluster_EW10} deploys cluster structure in order to implement common channel control, medium access  with multiple channel and channel allocation. 
The node with the maximum number of common channels within its k-hop neighborhood is chosen as cluster head, but how to avoid one node appearing in multiple clusters is not given consideration.

Clustering robustness is considered in~\cite{Lazos09, LIU_TMC11_2}.
The authors propose a distributed scheme where the metric is the product of cluster size and the number of common control channels.
This scheme involves both cluster size and number of CCCs, but it is inherently flawed.
With the metric, cluster could be formed only due to one factor of the two, e.g. a spectrum rich node will exclude its neighbour to form a cluster by itself.
%emphasis on improving the numbers of common channels within clusters, in order to strengthening robustness of clusters, but the perused metric is not examined or proved to be able to maintain cluster structure under vigorous intensity of primary users’ activities.
%The authors consider the balance between the number of idle common channels within cluster and cluster size and propose an algorithm that increases the number of common channels within clusters. 
%However, this work neglects the issue of connectivity between clusters. 
%One drawback of this scheme is, in order to increase the number of common channels within clusters, the scheme excludes certain CR nodes from the formed clusters, so that isolated nodes have to form clusters themselves. 
Besides, this scheme leads to a high variance on the size of clusters, which is not desired in certain applications as discussed in~\cite{clustering_globecom11, cluster_EW10}.

@@

We assume that in addition to the licensed channels, there is one dedicated control channel.
This control channel could be in ISM band or other reserved spectrum which is exclusively used for transmitting control messages.
Actually, the control messages involved in the clustering process can be transmitted on available licensed channels through a rendezvous process by channel hopping~\cite{channelHopping_Rendezvous_2014, Gu_distributed_rendezvous_2014}, i.e., two neighbouring nodes establish communication on the same channel.
Over the control channel, a secondary user $i$ can exchange its spectrum sensing result $K_i$ to any $i' \in \text{Nb}(i)$.
It is available for any secondary node $i$ to exchange control messages with any other node in its neighborhood $\text{Nb}(i)$ during the cluster formation phase.
$\text{Nb}(i)$ is simply defined as the set of nodes located within the transmission range of $i$.

@@

\subsection{Clustering}
\label{def_cluster}
%We give the description of cluster in the context of CRN, all the clusters discussed in this paper comply with it.
%A cluster $C$ is composed with one cluster head and cluster members, which satisfies the following conditions:
%\begin{itemize}
%\item Cluster head $H_C$ is able to communicate with any cluster member \textit{directly}, \ie for any cluster member $i\in C$, there is $i\in \texttt{Nb}_{H_C}$.
%\item There exists as least one common control channel in the cluster, \ie $\cap_{i\in C} V_i \neq \emptyset$.
%\end{itemize}
In this section, we describe what a cluster in the context of CRNs means.
A cluster $C \subseteq \mathcal{N}$ is a set of secondary nodes consisting of a cluster head $h_{\textsf{C}}$ and a number of cluster members.
The cluster head is able to communicate with any cluster member directly.
In other terms, for any cluster member $i \in C$, $i \in \text{Nb} (h_C) $ holds.

@@

\section{Clustering Algorithm: ROSS}
\label{ross}

In this section we introduce the distributed clustering scheme ROSS which leads to robust clusters against PU' influence.
ROSS consists of two cascaded phases: \textit{cluster formation} and \textit{membership clarification}.
With ROSS, CR nodes form clusters on the basis of the proximity of the available spectrum in their neighbourhood. Afterwards, CR nodes belong to one certain cluster. %by interacting with certain neighbours.
%is based on the spectrum sensing results in its neighbours , then

@@

\item \textit{Social connectivity degree} $g_i$: $g_i=|\bigcap_{j\in \text{Nb}(i)\cup i}K_j|$. It is the number of common channels available to $i$ and all its neighbors.
$g_i$ represents the ability of $i$ to form a robust cluster with its neighbours.
\end{itemize}
Individual connectivity degree $d_i$ and social connectivity degree $g_i$ together form the \textit{connectivity vector}.
Figure~\ref{fig1} illustrates an example CRN where each node's connectivity vector is calculated and shown.	
\begin{figure}[ht!]
  \centering
\includegraphics[width=0.7\linewidth]{figure1.pdf}
% \includegraphics{figure1.pdf}
	\caption{Connectivity graph and the connectivity vector $(d_i, g_i)$ for each node. The available channels sensed by each node are: $K_A=\{1,2,3,4,5,6,10\}, K_B=\{1,2,3,5,7\}, K_C=\{1,3,4,10\}, K_D=\{1,2,3,5\}, K_E=\{2,3,5,7\}, K_F=\{2,4,5,6,7\}, K_G=\{1,2,3,4,8\}, K_H=\{1,2,5,8\}$. Dashed edge indicates the end nodes are within each other's transmission range. The number of common channels between two nodes is shown by the edge.}
	\label{fig1}
\end{figure}

>>>>

\definecolor{slightgray}{gray}{.90} 
\usepackage{rotating}
\usepackage{hhline}
\usepackage{float}
\usepackage{caption}% http://ctan.org/pkg/caption
\captionsetup[table]{format=plain,labelformat=simple,labelsep=period}%

@@

\newcommand{\eg}{e.g., }
\newcommand{\ie}{i.e., }
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
%\usepackage{algorithmic}
\usepackage{times}
%\usepackage{mathptmx}
\usepackage{mathtools}
\usepackage[draft,nomargin,marginclue,footnote,silent]{fixme}
\setcounter{tocdepth}{2} %table of contents

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
Cluster structure in cognitive radio networks facilitates cooperative spectrum sensing, routing and other functionalities.
The unlicensed channels, which are available for every member of a group of cognitive radio users, consolidate the group into a cluster, and the availability of unlicensed channels decides the robustness of that cluster against the licensed users' influence.
%Thus there should be more unlicensed channels in the clusters.
This paper analyses the problem that how to form robust clusters in cognitive radio network, so that more cognitive radio users can get benefits from cluster structure even when the primary users' operation are intense.
%In the process of forming clusters, every secondary user decides with whom to form a cluster, or which cluster to join.
We provide a formal description of robust clustering problem, prove it to be NP-hard and propose a centralized solution, besides, a distributed solution is proposed to suit the dynamics in the ad hoc cognitive radio network.
Congestion game model is adopted to analyse the process of cluster formation, which not only contributes designing the distributed clustering scheme directly, but also provides the guarantee of convergence into Nash Equilibrium and convergence speed.
Our proposed clustering solution is versatile to fulfil some other requirements such as faster convergence and cluster size control.
The proposed distributed clustering scheme outperforms the related work in terms of cluster robustness, convergence speed and overhead.
%Besides, we prove the clustering problem is NP-hard, and also propose the centralized solution.
The extensive simulation supports our claims.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
cognitive radio, robust cluster, game theory, congestion game, distributed, centralised, cluster size control.
\end{IEEEkeywords}

@@

\section{Introduction}
\label{intro}
%CR concept
\IEEEPARstart{C}{ognitive} radio (CR) is a promising technology to solve the spectrum scarcity problem~\cite{Mitola}.
Licensed users access the spectrum allocated to them whenever there is information to be transmitted.
In contrast, unlicensed users can only access the licensed spectrum after validating the channel is unoccupied by licensed users, where spectrum sensing~\cite{sensing_survey_2009} plays an important role in this process.
%This refers to the process of sensing a particular channel and verifying (with a previously specified probability of error) it is not used by a primary user currently.
In this hierarchical spectrum access model~\cite{zhao_survey_DSA_2007}, the licensed users are also called primary users (PU), and the CR users are known as secondary users and constitute the cognitive radio networks (CRN).
%
%cluster
As to the operation of CRN, efficient spectrum sensing is identified to be critical to the success of cognitive radio networks~\cite{Sahai_FundamentalDesignTradeoffs2006}.
Cooperative spectrum sensing is able to effectively cope with noise uncertainty and channel fading, thus remarkably improving the sensing accuracy~\cite{coorperativeSensing_Akyildiz11}.
Collaborative sensing relies on the consensus of CR users within certain area, and decreases considerably the negative probability due to the random channel effects like fading and shadowing.
In this regard, clustering is regarded as an effective method in cooperative spectrum sensing~\cite{Sun07_clustering_spectrum_secsing, Zhao07}, as a cluster forms adjacent secondary users as a collectivity to perform spectrum sensing together.
%As a result, more accurate sensing result can be obtained by collaborative sensing, and the improvement on spectrum sensing decreases the interference originating from CR users to primary users, which is highly desirable.
%Also,  prevents CR users from using channels that are occupied by primary users. 
Clustering is also efficient to enable all CR devices within the same cluster to stop payload transmission on the operating channel and initiate the sensing process, so that all the CR users\footnote{The term $\textit{user}$ and $\textit{node}$ are used interchangeably in this paper, in particular, user is used when its networking or cognitive ability are discussed or stressed, and node is used when the network topology is discussed.} within the one cluster are able to vacate the channel swiftly when primary users are detected by at least one CR node residing in the cluster~\cite{willkomm08}.
With cluster structure, as CR users can be notified by cluster head (CH) or other cluster members about the possible collision, the possibility for them to interfere neighbouring clusters is reduced~\cite{centralizedSharing80222}. 
Clustering algorithm has also proposed to support routing in cognitive ad-hoc networks~\cite{Abbasi_survey_07}.

@@

%[robustness issue for clustering]
The communication within a cluster is conducted in the spectrum which is available for every member in that cluster.
Usually there are multiple unlicensed channels available for all the members in a cluster, which are referred as \textit{common control channels} (CCC).
When one or several members can not use one certain CCC because primary users are detected to appear on that channel, this channel will be excluded from the set of CCCs, in particular, if this channel is the working channel, then all the cluster members switch to another channel in the set of CCCs.
In the context of CRN, as the activity of primary users is controlled by licensed operators which are generally not known to CR users, the availability of CCCs for the formed clusters  is totally decided by primary users' activity.
In other words, the availability of CCCs for clusters is passive and can not guaranteed.
%
In CRN, one cluster survives the influence of primary users when at least one CCC is available in that cluster.
As the primary users' operation is assumed to be unknown to the CR users, a cluster formed with more CCCs will survive with higher probability.
Thus the number of CCCs in one cluster indicates the robustness of it when facing the unpredictable influence from primary users.
%It is easy to see that forming clusters with different neighbours leads to different amount of CCCs in the clusters.
%As a result, how to form the clusters plays an important role on the robustness of clusters in CRN.

@@

To solely pursue cluster robustness against the primary users' activity, \ie to achieve more common channels within clusters, the ultimately best clustering strategy is ironically that every node constitutes one cluster (called as \textit{singleton cluster} in this paper).
%In that case, the common channels within cluster are the available channels available at that node's place.
Apparently this contradicts our motivation of proposing cluster in cognitive radio network to enable cooperative decision making.
This contradiction indicates that, the robustness discussed in terms of number of common channels carries little meaning when the sizes of formed clusters are not given consideration.
Besides, cluster size plays import roles in certain aspects.
For instance, cluster size is one decisive factor in power preservation~\cite{clustering_globecom11, EnergyEfficientClusteringRouting_2015}, and it also influences the accuracy of cooperative spectrum sensing~\cite{Consensus_based_clustering12}.
%thesis
%Hence, cluster size should be given consideration when discussing cluster robustness against primary users.

In this paper, a decentralized clustering approach ROSS (RObust Spectrum Sharing) is proposed to cover the issues of robustness and size control of clusters in CRN.
ROSS is able to form clusters with desired sizes, and when compared with previous works, ROSS involves smaller signalling overhead, and the generated clusters are significantly more robust against the primary users which appear after the clusters are formed.
%thesis
%and the generated clusters are more robust than the other clustering scheme which has claims on cluster robustness, \ie more secondary users residing in clusters against increasing influence from primary users.
%
%ROSS selects cluster head through coordination within its neighborhood, and then cluster membership is decided locally and its fast convergence is proved under game theoretic framework. 
%For our scheme we can prove convergence in cluster formation phase and resolve ambiguities with respect to cluster membership in a game-theoretic setting. 
We also propose the light weighted versions of ROSS, which involve less overheads and thus are more suitable for fast deployment.
Throughout this paper, we refer the clustering schemes on the basis of ROSS as \textit{variants of ROSS}, \ie the fast versions, or that with size control feature.

The rest of paper is organized as follows. 
After reviewing related work in section~\ref{related_work}, we present the system model in Section~\ref{sec:model}. 
In Section~\ref{problem}, problem formulation is presented and the problem is thoroughly analysed.
The centralized clustering scheme and the our proposed clustering scheme ROSS are introduced in Section~\ref{centralized_opt} and \ref{ross} respectively.
%The clustering problem is given through analysis and a centralized scheme is proposed in section~\ref{centralized_scheme}.
Extensive performance evaluation is in section~\ref{performance}.
Finally, we conclude our work and point out direction future research in section~\ref{conclusion}.

@@

Prior to the emergence of cognitive radio, clustering has been proposed as an important method to manage network in ad hoc networks~\cite{Kawadia03,Lin97adaptiveclustering,Basagni99}, wireless mesh networks and sensor networks~\cite{Abbasi_survey_07}. 
In ad hoc and mesh networks, the major focus of clustering is to preserve connectivity (under static channel conditions) or to improve routing.
In sensor networks, the emphasis of clustering is on longevity and coverage.
Overhead generated by clustering in ad hoc network is analysed in~\cite{clusterRoutingOverhead02infocom, clusterRoutingOverhead_wcnc04}.

As to cognitive radio networks, clustering schemes are also proposed, which target different aspects.
Work~\cite{Consensus_based_clustering12} improves spectrum sensing ability by grouping the CR users
with potentially best detection performance into the same cluster.
Clustering scheme~\cite{clustering_globecom11} obtains the best cluster size which minimizes power consumption caused by communication within and among clusters.
\cite{clustering_globecom11} proposes clustering strategy in cognitive radio network, which looks into the relationship between cluster size and power consumption and accordingly controlling the cluster size to decrease power consumption.
%The works~\cite{} introduced in Introduction section don't provide solution that how are the clusters formed.
%There have been several clustering schemes tailored for CRNs. 
Cogmesh is proposed in~\cite{Chen07} to construct clusters with the neighbouring nodes which share local common channels.
This method pays attention to the communication among the formed clusters, but cluster robustness is not considered.
\cite{TWC2012_cooperative_communication} targets on the QoS poisoning and energy efficiency. 
This approach first decides on the relay nodes which minimize transmission power consumption, then the chosen nodes become cluster heads and clusters are formed in a dynamic coalition process.
This work emphasis on power efficiency and doesn't take into account the channel availability and the issue of robustness of the formed clusters.
In~\cite{Zhao07, Affinity_clustering_09icccn}, the channel available to the largest set of one-hop neighbours is selected as common channel which yields a partition of the CRN into clusters. 
This approach minimizes the set of distinct frequency bands (and hence, the set of clusters) used as common channels within the CRN.
However, bigger cluster sizes generally lead to less options within one cluster to switch to if the common channel is reclaimed by a primary node. 
Hence, this scheme does not provide robustness to formed clusters. 
\cite{cluster_EW10} deploys cluster structure in order to implement common channel control, medium access  with multiple channel and channel allocation. 
The node with the maximum number of common channels within its k-hop neighborhood is chosen as cluster head, but how to avoid one node appearing in multiple clusters is not given consideration.
A event-driven clustering scheme is proposed for cognitive radio sensor network in \cite{Ozger_cluster_crsn_13}.
The cluster heads are selected according to node degree, spectrum availability and the distance to the sink, and the clusters are no longer available after the event ends.

Clustering robustness is considered in~\cite{Lazos09, LIU_TMC11_2, mansoor_15_cluster_robust}.
The first two propose a distributed scheme where the metric is the product of cluster size and the number of common control channels.
This scheme involves both cluster size and the number of CCCs, but it is inherently flawed.
With the metric, cluster could be formed only due to one factor of the two, e.g. a spectrum rich node will exclude its neighbour to form a cluster by itself.
%emphasis on improving the numbers of common channels within clusters, in order to strengthening robustness of clusters, but the perused metric is not examined or proved to be able to maintain cluster structure under vigorous intensity of primary users’ activities.
%The authors consider the balance between the number of idle common channels within cluster and cluster size and propose an algorithm that increases the number of common channels within clusters. 
%However, this work neglects the issue of connectivity between clusters. 
%One drawback of this scheme is, in order to increase the number of common channels within clusters, the scheme excludes certain CR nodes from the formed clusters, so that isolated nodes have to form clusters themselves. 
Besides, this scheme leads to a high variance on the size of clusters, which is not desired in certain applications as discussed in~\cite{clustering_globecom11, cluster_EW10}.
\cite{mansoor_15_cluster_robust} presents a heuristic method to form clusters, although the authors claim robustness is one goal to achieve, they actually uphold the minimum number of formed clusters.

@@

We assume that in addition to the licensed channels, there is one dedicated control channel.
This control channel could be in ISM band or other reserved spectrum which is exclusively used for transmitting control messages.
Actually, the control messages involved in the clustering process can also be transmitted on the available licensed channels through a rendezvous process by channel hopping~\cite{channelHopping_Rendezvous_2014, Gu_distributed_rendezvous_2014}, i.e., two neighbouring nodes establish communication on the same channel.
Over the control channel, a secondary user $i$ can exchange its spectrum sensing result $K_i$ to any $i' \in \text{Nb}(i)$.
It is available for any secondary node $i$ to exchange control messages with any other node in its neighborhood $\text{Nb}(i)$ during the cluster formation phase.
$\text{Nb}(i)$ is simply defined as the set of nodes located within the transmission range of $i$.

@@

\section{Robust Clustering Problem in CRN}
\label{problem}
In this section, we describe the definition of cluster in the context of CRNs, and the problem of robust clustering.

\subsection{Cluster in CRN}
\label{def_cluster}
%We give the description of cluster in the context of CRN, all the clusters discussed in this paper comply with it.
%A cluster $C$ is composed with one cluster head and cluster members, which satisfies the following conditions:
%\begin{itemize}
%\item Cluster head $H_C$ is able to communicate with any cluster member \textit{directly}, \ie for any cluster member $i\in C$, there is $i\in \texttt{Nb}_{H_C}$.
%\item There exists as least one common control channel in the cluster, \ie $\cap_{i\in C} V_i \neq \emptyset$.
%\end{itemize}
A cluster $C \subseteq \mathcal{N}$ is a set of secondary nodes consisting of a cluster head $h_C$ and a number of cluster members.
The cluster head is able to communicate with any cluster member directly.
In other terms, for any cluster member $i \in C$, $i \in \text{Nb} (h_C) $ holds.

@@

\subsection{Robust Clustering Problem}
\label{centralized_scheme}

The centralized clustering scheme aims to form clusters with desired sizes, meanwhile the total number of CCCs of all clusters is maximized.
In the following, we refer this problem as \textit{centralized clustering}, and give the formal problem definition.

\begin{mydef}
\label{def_centralized_clustering}
\textit{Robust clustering problem in CRN.}

%The decision version of \textit{weighted exact cover problem}: 
%Given an universe $U$, and collection $S=\{s_1, s_2, \ldots, s_m\}$ where each subset $s_i\subseteq U$ and is given a weight $w_i$, whether there exists a collection of subsets $\mathcal{C}$ and constant number $\lambda$, so that the union of $\mathcal{C}$ equals to $U$, $s_j\cap s_{j'} = \emptyset$ for different $j$ and $j'\in \{1,2,\ldots, m\}$, and $\sum_{i\in J} w_i \geq \lambda$.

\subsection{Complexity of the Robust Clustering Problem}
In this section we investigate the complexity of robust clustering problem.
Theorem~\ref{theorem1} tells centralized clustering problem in CRN is one NP-hard problem.

\begin{theorem}
\label{theorem1}
CRN clustering problem is NP-hard, when the maximum size of clusters $k\geqslant 3$.
%Assume a CRN can be represented by a connected graph, and there is at least one common channel between any pair of neighbours, then forming at least two CR nodes into one cluster is NP-complete.
\end{theorem}
The proof is in Appendix~\ref{proof_theorem1}.
In the following sections, both centralized and distributed solutions will be introduced.

\section{Centralized Solution for Robust Clustering}
\label{centralized_opt}
%Exact cover problem can be solved with Knuth's Algorithm X~\cite{dancingLinks_Knuth} as it finds out all the instances of exact cover, then we can choose the one with the biggest sum of weights. 
As the robust clustering problem in CRN is NP hard, there is no polynomial algorithm to solve the problem, \ie maximizing the sum of CCCs in all non-singleton clusters.
Thus, we formulate the problem into a binary linear programming problem, where the objective function and the constraints are heuristic, in particular, the objective is different from the object in the Definition~\ref{def_centralized_clustering}.
%Note that binary linear programming is in NP-complete.

Given a CRN $\mathcal{N}$ and desired cluster size $\delta$, we obtain a collection of clusters $\mathcal{G}$ which contains all the \textit{legitimate} clusters, and the sizes of these clusters are $1,2,\ldots,\delta$.
Legitimate clusters are the clusters which satisfy the conditions in Section~\ref{def_cluster}. 
Note that the legitimate clusters include the \textit{singleton} ones, \ie the cluster which has only one CR node, so that we can guarantee the partition of any network is always feasible.
%
With $N=|\mathcal{N}|, G=|\mathcal{G}|$, we construct a constant $G\times N$ matrix $Q_{G\text{x}N}$. 
The element of matrix $Q$ is $q_{ij}$, where the subscript $i$ is the index of legitimate cluster, and $j$ is the node ID of one CR node.
%Note that $C_i$ means $i$th cluster in the collection $\mathcal{G}$, doesn't denote the cluster where cluster head is $i$, this notation is only valid in this subsection.
There are $i\in \{1,2,\cdots,G-1, G\}$, and $j\in \{1,2,\cdots,N-1, N\}$.
Element $q_{ij}= |K(C_i)|$ if node $j\in C_i$, and $q_{ij}= 0$ if $j\notin C_i$.
In other words, each non-zero element $q_{ij}$ denotes the number of CCCs of the cluster $i$ where node $j$ resides.

The objective is a sum of two components, the first component is the sum of products of cluster size and the corresponding number of CCCs, which is the sole metric adopted by the scheme SOC~\cite{Lazos09}.
The second component is the \textit{punishment} for choosing the clusters whose sizes are not $\delta$.
In fact, the second part is particularly designed to eliminates the drawbacks of SOC, \ie SOC produces a large number of singleton clusters and a few large clusters.
In practical computation, we minimize the opposite of the first part, and the punishment is a positive value.
%
The first constraint restricts each node $j$ to reside in exactly one cluster.
In the second constraint, $w_i$ is an auxiliary binary variable which denotes whether cluster $C_i$ is chosen by the solution, in particular, 
$$
w_i = \left\{ \begin{array}{rl}
0 &\mbox{if $i$th legitimate cluster $C_i$ is chosen} \\
1 &\mbox{if $i$th legitimate cluster $C_i$ is not chosen} \\
\end{array} \right.
$$
The second constraint regulates that when the $i$th legitimate cluster $C_i$ is chosen, the number of elements which equal to 1 in the $i$th row of the matrix $X$ is $|C_i|$.
%
Now we explain how does the mechanism of the punishment in the objective work. 
The parameter $p$ is defined as follows,
$$
p = \left\{ \begin{array}{rl}
0 &\mbox{ if $|C_i|=\delta$} \\
\alpha_1 &\mbox{if $|C_i|=\delta-1$} \\
\alpha_2 &\mbox{if $|C_i|=\delta-2$} \\
\dots
\end{array} \right.
$$
where $\alpha_i>0$ and increases when $|C_i|$ diverges from $\delta$.
Because of $w_i$, any chosen cluster ($w=0$) brings certain \textit{punishment}.
When the chosen cluster's size is desired size $\delta$, the punishment is zero.
In contrary, when the chosen cluster's size diverges from $\delta$, the objective function suffers \textit{loss}.
In particular, when $w_i=0$ and $|C_i|=1$, the punishment is the most severe.
This design doesn't follow the definition of $f(C)$ in Definition~\ref{def_centralized_clustering} strictly, where $f(C_i)=0$ when $|C_i|=1$, but our design echoes the definition by exerting the most severe punishment on the singleton clusters in the clustering solution.
Choice of $\alpha_i$ affects the resultant clusters.

%Note that the proposed centralized solution is heuristic.
%We reiterate the reasons for pursuing the heuristic scheme, 
%Note the collection of legitimate clusters is dependant on the network topology and spectrum availability in the network, thus to each specific CRN, the space of solution is different.

\section{Distributed Clustering Algorithm: ROSS}
\label{ross}
The cluster formation of one CRN can be obtained by solving the proposed optimization problem, but centralized scheme is inherently not suitable for the CRN which involves dynamics, \eg the mobility of CR users and the variance of the available spectrum due to primary users' operation.
In this section we introduce the distributed clustering scheme ROSS.
%
ROSS consists of two cascaded phases: \textit{cluster formation} and \textit{membership clarification}.
With ROSS, CR nodes form clusters based on the proximity of the available spectrum in their neighbourhood. Afterwards, CR nodes belong to one certain cluster. %by interacting with certain neighbours.
%is based on the spectrum sensing results in its neighbours , then

@@

\item \textit{Social connectivity degree} $g_i$: $g_i=|\bigcap_{j\in \text{Nb}(i)\cup i}K_j|$. It is the number of common channels available to $i$ and all its neighbors.
$g_i$ represents the ability of $i$ to form a robust cluster with its neighbours.
\end{itemize}
Individual connectivity degree $d_i$ and social connectivity degree $g_i$ together form the \textit{connectivity vector}.
Figure~\ref{fig1} illustrates an example CRN where each node's connectivity vector is shown.	
\begin{figure}[ht!]
  \centering
\includegraphics[width=0.7\linewidth]{figure1.pdf}
% \includegraphics{figure1.pdf}
	\caption{Connectivity graph of the example CRN and the connectivity vector $(d_i, g_i)$ for each node. The sets of the indices of the available channels sensed by each node are: $K_A=\{1,2,3,4,5,6,10\}, K_B=\{1,2,3,5,7\}, K_C=\{1,3,4,10\}, K_D=\{1,2,3,5\}, K_E=\{2,3,5,7\}, K_F=\{2,4,5,6,7\}, K_G=\{1,2,3,4,8\}, K_H=\{1,2,5,8\}$. Dashed edge indicates the end nodes are within each other's transmission range.}
	% The number of common channels between two nodes is shown by the edge.
	\label{fig1}
\end{figure}
